---
title: "Getting Started with R"
author: "asif zubair"
date: "February 29, 2016"
output: html_document
---

R package repositories
- [CRAN](https://cran.r-project.org)  
    - [Statistical Genetics](https://cran.r-project.org/web/views/Genetics.html)  
- [Bioconductor](http://bioconductor.org)  

A fun way for learning the ropes with R is using the package `swirl`

```{r}
install.packages("swirl")
library(swirl)
```

This will open a dialogue in the console and start teaching you `R` basics and other cool stuff.

R programming basics:
- Data Types
- Assignment
- Control Flow
    - Loops
    - Conditionals
- Functions

__Exercises:__

- what is the version of R you are using ?
```
> version
               _                           
platform       x86_64-apple-darwin13.4.0   
arch           x86_64                      
os             darwin13.4.0                
system         x86_64, darwin13.4.0        
status                                     
major          3                           
minor          2.0                         
year           2015                        
month          04                          
day            16                          
svn rev        68180                       
language       R                           
version.string R version 3.2.0 (2015-04-16)
nickname       Full of Ingredients         
```


- R data types by way of `state` data
- please eneter `?state` on your console
```
state.area                #a NUMERIC vector giving the area of US states, in square miles
state.name                #a CHARACTER vector (note the quote marks) of state names 
sq.km <- state.area*2.59  #Arithmetic works on numeric vectors, e.g. convert sq miles to sq km
sq.km                     #... the new vector has the calculation applied to each element in turn
sqrt(sq.km)               #Many mathematical functions also apply to each element in turn 
range(state.area)         #But some functions return different length vectors (here, just the max & min).
length(state.area)        #and some, like this useful one, just return a single value.
```

- Use a `for` loop to determine the sum of squares of the first 25 natural numbers ?

In fact, let's try to solve this question using three methods . 

First the `for` loop:
```{r}
sum2 = 0
for (i in seq(25)){
    sum2 = sum2 + i*i
    }
print(sum2)
```

Great, but can we make this into a `function`:

```{r}
sum2<-function(n){
    temp = 0
    for (i in seq(n)){
    temp = temp + i*i
    }
    return(temp)
}
```

Cool! Let's use a mathematical identity:

```{r}
sum2 <- function(...){
    temp = (n*(n+1)*(2*n+1)/6)
    return(...)
}
```

and finally, let's do `recursion`.

```{r}
sum2<-function(n){
    if (n==1){
        return(1)
    } else {
        return(n*n + sum2(n-1))
    }
}
```

- The `cars` dataset is available in base R. You can type `cars` to see it. Use the `class` function to determine what type of object is cars.

```
cars                 package:datasets                  R Documentation
Speed and Stopping Distances of Cars
Description:
     The data give the speed of cars and the distances taken to stop.
     Note that the data were recorded in the 1920s.
Usage:
     cars
Format:
     A data frame with 50 observations on 2 variables.
       [,1]  speed  numeric  Speed (mph)            
       [,2]  dist   numeric  Stopping distance (ft) 
Source:
     Ezekiel, M. (1930) _Methods of Correlation Analysis_.  Wiley.
References:
     McNeil, D. R. (1977) _Interactive Data Analysis_.  Wiley.
```

- How many rows does the `cars` object have?
- What is the name of the second column of `cars`?
- The simplest way to extract the columns of a `matrix` or `data.frame` is using `[`. For example you can access the second column with `cars[,2]`. What is the average distance traveled in this dataset?
```{r}

```
- Familiarize yourself with the `which` function. What row of `cars` has a a distance of 85?

a linear regression example:

- Lets make some data
```{r}
x = seq(10)
y = 1 + 2*x + rnorm(10)
data = data.frame(x=x, y=y)
```

- let's see dependence of the data
```{r}
lm.xy <- lm(y ~ x, data=data)
summary(lm.xy)
```

- finally, let's see what we have computed
```{r}
plot(x, y, main="y vs x", pch=16)
abline(lm.xy, col="red", lwd=3)
```

Resources:
- [DataCamp](https://www.datacamp.com/home)'s free R courses  
- [Lynda](http://itservices.usc.edu/lynda) - use USC login  
- [r-bloggers](http://www.r-bloggers.com) - R tutorials and tricks  
- [stackoverflow](http://stackoverflow.com/questions/tagged/r) for troubleshooting  
